Ages of Stars and Planets Observed by Kepler

### 2024-04-23 paper/ms.tex
update ms

%%%%%%%%%%%
% LITHIUM %
%%%%%%%%%%%
% number of KOIs that are not known FPs that have a HIRES spectrum
\newcommand{\nkoiswithhires}{{1{,}500}}
% number of KOIs for which this work and Berger2018 overlap
\newcommand{\nbergeroverlap}{{500}}

@@

@@

We measured the lithium equivalent widths following a procedure
adapted from previous work \citep{Bouma_2021}.  Our stars of interest
are FGK stars, and so the continuum in the vicinity of the \ion{Li}{1}
6708\,\AA\ doublet is well-defined.  We Doppler corrected the spectra
to a common reference wavelength by cross-correlating against a high
S/N template for Kepler-1698, chosen because $\gamma$$\approx$9\kms,
$T_{\rm eff}$$\approx$5000\,K and $v\sin i$$\approx$5\,\kms, which
puts it in the middle of our sample's temperature range and gives it
mild line-broadening.  We then trimmed the Doppler-corrected spectra
to a local window centered on the lithium line, using a window width
of 15\,\AA (we considered 10\,\AA\ and 20\,\AA; the results were
broadly consistent).  We then continuum-normalized by fitting a
third-degree Chebyshev series, while excluding any regions that showed
absorption lines.  We then numerically integrated the resulting
spectrum using a one-component Gaussian with free amplitude, width,
and mean, and estimated uncertainties on the line width through a
Monte Carlo procedure that bootstraped against the local scatter in
the continuum.

%    %FIXME FIXME TODO LEFT OFF HERE
%    %FIXME FIXME TODO LEFT OFF HERE
%    %FIXME FIXME TODO LEFT OFF HERE
Our EW measurement approach does not correct for the neighboring
\ion{Fe}{1} 6707.44\,\AA\ blend.
To evaluate the accuracy and precision of our equivalent measurement
technique, after applying an initial iteration on the
\nplhoststarwgyroagewithgrazingandhighruwe\ aforementioned stars,
we compared our lithium equivalent widths
with those reported by \citet{2018ApJ...855..115B}.
For the \nbergeroverlap\ stars in both samples, we found {\bf X, Y,
Z.}

%    These HIRES spectra are available for \nkoiswithhires\ of the
%    \nkoisnofp\ non-false-positive KOIs.
%    In the broader stellar sample, they are available for a much smaller 
%    fraction of the stars observed by Kepler, because they were generally only
%    acquired due to the presence of a transiting planet candidate.
%    Given this selection function,
%    we leverage these spectra in our analysis as a way to scrutinize the
%    ages of just the planet sample.

@@

We calculate gyrochrone ages using \texttt{gyro-interp}
\citep{Bouma_2023}.  We designed this method to address
the issue that, especially over the first gigayear of stellar
spin-down, stars with the same mass and same rotation period can have
very different ages \citep[e.g.][]{Curtis_2019_ngc6811}.  The
gyrochrone age posterior should therefore incorporate this intrinsic
population-level scatter into the precision with which the age can be
inferred.  The three main assumptions behind the \texttt{gyro-interp}
framework are as follows.  {\it 1)} The stars are near the main
sequence, {\it 2)} their spindown is not influenced by binary
companions, and {\it 3)} the rotation period distribution for field
stars of a given mass and age is identical to that of equally aged
open clusters (metallicity differences, for instance, are ignored).  
Regarding the former two assumptions: pre-main-sequence contraction and
post-main-sequence expansion influence surface stellar rotation periods
in ways that we do not model.  Stars in e.g.~ close binaries might be
subject to tidal effects that we do not treat.

Using our adopted effective temperatures and rotation periods, we ran
\texttt{gyro-interp} over an age grid of 0 to 4\,Gyr, linearly spaced
over 500 grid points.  \texttt{gyro-interp} itself is calibrated
out to 4\,Gyr, the adopted age of the M67 cluster
\citep[see][]{2022ApJ...938..118D,Gruner_2023}.  To estimate the stellar
spin-down rate as a function of time and stellar temperature, this
framework interpolates between reference open clusters using smooth
cubic hermite polynomials.  We calculate the probability of the
rotation-based age $t_{\rm gyro}$, given the observed rotation
periods, effective temperatures, and their uncertainties by
integrating Equation~1 of \citet{Bouma_2023} at the default grid
resolution.  This procedure marginalizes over the intrinsic age and
temperature dependent scatter that is observed in the cluster
sequences to yield statistical age uncertainties.

@@

\subsection{Gyrochronology Quality Flags}
\label{subsec:flags}
We calculated gyrochrone age posteriors for all \nuniqstarsantosrot\
stars with reported rotation periods and effective temperatures.  To
flag stars for which we suspect these ages may not in fact be valid,
we then build a set of bit-quality flags.  When and how these flags should
be applied depends on the question being asked.  If the goal is to
construct a false-positive free sample, they might all be applied.  If
the goal is constructing a complete sample of young stars, then
consider the examples of Kepler~1627Ab ($\approx$40\,Myr) and
Kepler~51c ($\approx$625\,Myr).  The former has a high RUWE due to a
resolved binary companion \citep{Bouma_2022a}; the latter is on a
grazing orbit \citep{2014ApJ...783...53M}.  Our quality flags enable
selecting for or against such cases by selecting from the following
bits.

@@

{\it Surface gravity (bit 1)}---We flagged stars as possible subgiants
if they had $\log g < 4.2$.  We set this threshold by examining the
relevant flag from \citet{berger_2018_radii_evolnstates},
cross-matched against \citet{Berger_2020a_catalog}.
%\citet{berger_2018_radii_evolnstates} performed this classification
%using the PARSEC evolutionary models, stellar radii, and DR25 stellar
%properties catalog effective temperatures \citep{Mathur_2017}.

@@

An important decision for our subsequent analysis is how to combine
the above flags in order to decide whether or not a star is suited for
gyrochronology.  By default, we assume that a star is suitable for
gyrochronology if none of bits zero through nine are raised.

\subsection{Planet Quality Flags}
\label{subsec:plflags}
For the subset of our analysis that focuses on the properties of
planets, we use the following additional quality flags.

% see Erik's 2022 discussion; might want to shift to include a cut on
% disposition score.
{\it Candidate reliability}---no false positives; only ``planet
candidates'' and ``confirmed planets''.

{\it Candidate S/N}---the false positive rate increases greatly toward
Kepler's noise floor for planet detection (CITE).  We required a S/N
in excess of the usual floor, through a cut on the ``multiple event
statistic'' (MES): we required ${\rm MES}$$>$10.

{\it Grazing planets}---grazing objects ($b$>0.8) often provide
non-robust planetary sizes due to the size-impact parameter
degeneracy.  Particularly for large planets, they can also include
astrophysical false positives at higher rates (due to the same
degeneracy).  While we include a flag for such objects, we by default
do not exclude them from our analysis; this is because...{\bf justify}.

@@

\begin{figure}[!t]
	\begin{center}
		\leavevmode
		\includegraphics[width=0.49\textwidth]{li_vs_teff_koi_X_S19S21dquality_eagles_logy.pdf}
	\end{center}
	\vspace{-0.6cm}
	\caption{
    {\bf Equivalent widths (EW) of the \ion{Li}{1} 6708\,\AA\ doublet
    from Keck/HIRES spectra collected between 2009 and 2024.} Lines
    are the ``mean'' isochrones from the same models as in
    Figure~\ref{fig:models}.  The logarithmic scale highlights the
    dynamica range of both detections and upper limits.
		\label{fig:li_vs_teff}
	}
\end{figure}

To calculate lithium ages, we use a combination of \texttt{EAGLES}
\citep{Jeffries_2023} and \texttt{BAFFLES}
\citep{Stanford-Moore_2020}.
The former tool was calibrated on stars observed by the Gaia-ESO
spectroscopic survey in 52 open clusters, with ages spanning 2 to
6{,}000\,Myr.  
Example \texttt{EAGLES} isochrones, as well as the modelled intrinsic dispersion
about those isochrones, are shown in Figure~\ref{fig:li_vs_teff}.
While our preference would have been to exclusively use
\texttt{EAGLES} due to its homogeneous analysis approach ...

(Mention Linear prior in age.)

@@

In terms of detected planet counts, our rotation-based ages for
the Kepler sample yield \{\nplyounggyro, \nplmidgyro, \nploldgyro\}
detected planets in the 0-1\,Gyr, 1-2\,Gyr, and 2-3\,Gyr bins.
For the sub-gigayear bin in particular, which holds specific weight in
studies of planet evolution, while there are \nplyounggyro\ planets
with median ages below 1\,Gyr, requiring $t_{\rm
gyro}$$<$1\,Gyr at 2$\sigma$ confidence yields
\nplyounggyrotwosigma\ planets orbiting \nplhostsyounggyrotwosigma\
stars.
The stellar and planetary properties of this latter sample of young
planets are given in Table~\ref{tab:youngplanets}.

@@

\begin{figure*}[!t]
	\begin{center}
		\subfloat{
			\includegraphics[width=0.78\textwidth]{rp_vs_period_scatter_20240415_colorbyage_showaux-gyro_selsnrupper3_strictyoung.pdf}
		}

		\subfloat{
			\includegraphics[width=0.78\textwidth]{rp_vs_period_scatter_20240415_colorbyage_showaux-gyro_anyyoung_anyyoung.pdf}
		}
	\end{center}
	\vspace{-0.5cm}

	\caption{
		{\bf Sizes, orbital periods, and rotation-based ages of transiting exoplanets younger than one billion years.} 
		Two possible selection functions for what defines a young, sub-gigayear, planet
		are imposed on the {\it top} and {\it bottom}: the former requires $t$<1\,Gyr with
		$t/\sigma_{\rm t}$>3, the latter also allows planets younger than 1\,Gyr at 2$\sigma$.
		Circles are previously reported measurements from the literature.
		Stars are newly calculated ages for Kepler planets from this work; circles
		are planets from the literature meeting the same requirements.
		Gray points are transiting planets from the NASA Exoplanet Archive older than
		1\,Gyr.
		The three age bins span 10-46\,Myr, 46-215\,Myr, and 215-1000\,Myr.
		\label{fig:rp_period_age_results}
	}
\end{figure*}

@@

\startlongtable
\begin{deluxetable*}{lllllllrrrrrr}
	%\tabletypesize{\footnotesize}
	\tabletypesize{\scriptsize}
	\tablecaption{Ages of Kepler planets and planet candidates.  Rows are sorted by the minimum of either the rotation or lithium-based age; this version of the table is truncated to planets younger than one billion years.  ``Confirmed'' planets appear before``candidate'' planets.
		The full machine-readable table contains ages and age limits for \nnonfopkoissomeageinfo\ non-false positive KOIs with MES$>$10.
		The bit quality flags for the rotation-based ages, $Q_{\rm gyro}$, are described in Section~\ref{subsec:flags}.
		%The non-truncated machine readable versions are accessible both
		%through the online journal, and at
		%\url{https://zenodo.org/record/8327508}.
		\label{tab:youngplanets}}
	%\toprule
	%\midrule
	%\endhead
	\startdata
  KOI & Kepler &  $T_{\rm eff}$ & $P_{\rm rot}$ & EW$_{\rm Li}$ & $t_{\rm gyro}$ & $t_{\rm Li} $ & $R_{\rm p}$ & $P_{\rm orb}$ & $f_{\rm RUWE}$ & $f_{\rm grazing}$ & $Q_{\rm gyro}$ &  Spec? \\
  -- &   -- & K & days & m\AA & Myr & Myr &    Earths &    days &      bool &   bool  & int & bool \\
	\hline
	\input{table_subgyr_confirmed} %\\
	%\hline
	\input{table_subgyr_candidate}
	\enddata
%	\tablecomments{This table includes X, Y, Z... The machine-readable version,
%		available online, includes additional columns for .... 
%	}
\end{deluxetable*}
%\endlongtable

>>>>

%%%%%%%%%%%
% LITHIUM %
%%%%%%%%%%%
% number of KOIs that are not known FPs that have a HIRES spectrum
%\newcommand{\nkoiswithhires}{{1{,}500}}
% number of KOIs for which this work and Berger2018 overlap
%\newcommand{\nbergeroverlap}{{500}}

@@

Owing to the success of the transit and radial velocity techniques,
exoplanet science is going through an age of abundance in which we are
beginning to explore thousands of new worlds that orbit in close
proximity to their host stars.
However, most of these exoplanets are old, with ages comparable to
that of the solar system.
This limits our ability to understand their origins, and in particular
to study processes such as dynamical migration (CITE), thermal cooling
(CITE), and atmospheric loss (CITE) that typically happen soon after
disk dispersal.
Finding and characterizing young ($<$1\,Gyr) exoplanets is one way to
advance our understanding of how and when these processes occur.

@@

We measured the lithium equivalent widths using a procedure
adapted from previous work \citep{Bouma_2021}.  Our stars of interest
are FGK stars, and so the continuum in the vicinity of the \ion{Li}{1}
6708\,\AA\ doublet is well-defined.  We Doppler corrected the spectra
to a common reference wavelength by cross-correlating against a high
S/N template for Kepler-1698, chosen because $\gamma$$\approx$9\kms,
$T_{\rm eff}$$\approx$5000\,K and $v\sin i$$\approx$5\,\kms, which
puts it in the middle of our sample's temperature range and gives it
mild line-broadening.  We then trimmed the Doppler-corrected spectra
to a local window centered on the lithium line, using a window width
of 15\,\AA (we considered 10\,\AA\ and 20\,\AA; the results were
broadly consistent).  We continuum-normalized by fitting a
third-degree Chebyshev series, while excluding regions with
absorption lines.  We then numerically integrated the resulting
spectrum using a one-component Gaussian with free amplitude, width,
and mean, and estimated uncertainties on the line width through a
Monte Carlo procedure that bootstraped against the local scatter in
the continuum.

Our EW measurement approach does not correct for the neighboring
\ion{Fe}{1} 6707.44\,\AA\ blend.  To evaluate the accuracy and
precision of our equivalent measurement technique, after applying an
initial iteration on the \nlithiumstars\ KOIs with spectra, we
compared our lithium equivalent widths with those reported by
\citet{2018ApJ...855..115B}.  For the \nbergeroverlap\ stars in both
samples, we found broad agreement at $>$30\,m\AA, and significant
differences at $<$30\,m\AA because \citet{2018ApJ...855..115B}
required positive EWs, while we allow for negative ones.  At
$>$30\,m\AA, there is however a small offset in our respective
scales, such that (B24-B18)$=$7.50$\pm$8.55\,m\AA.  This offset is
caused by our non-treatment of the iron blend.  We therefore directly
subtract the constant mean value noted above when calculating
lithium-based ages.

@@

We calculate gyrochrone ages using \texttt{gyro-interp}
\citep{Bouma_2023}, which is a method designed to address the issue
that, especially over the first gigayear of stellar spin-down, stars
with the same mass and same rotation period can have very different
ages \citep[e.g.][]{Curtis_2019_ngc6811}.  The gyrochrone age
posterior should therefore incorporate this intrinsic population-level
scatter into the precision with which the age can be inferred.
Figure~\ref{fig:models} exemplifies this problem, particularly in
regions where the 0.1-1\,Gyr stars overlap.

The three main assumptions required for a rotation-based age dating
method like \texttt{gyro-interp} to be valid are as follows.  {\it 1)}
The target star must be near the main sequence, {\it 2)} its spindown
must not be influenced by binary companions, and {\it 3)} the rotation
period distribution for field stars of a given mass and age must be
identical to that of equally aged open clusters (metallicity
differences, for instance, are ignored).  Regarding the former two
assumptions: pre-main-sequence contraction and post-main-sequence
expansion influence surface stellar rotation periods in ways that we
do not model.  Stars in e.g.~ close binaries might be subject to tidal
effects that we do not treat.

Using our adopted effective temperatures and rotation periods, we ran
\texttt{gyro-interp} over an age grid of 0 to 4\,Gyr, linearly spaced
over 500 grid points.  \texttt{gyro-interp} itself is calibrated
out to 4\,Gyr, the adopted age of the M67 cluster
\citep[see][]{2022ApJ...938..118D,Gruner_2023}.  To estimate the stellar
spin-down rate as a function of time and stellar temperature, the
framework interpolates between reference open clusters using smooth
cubic hermite polynomials.  We calculate the probability of the
rotation-based age $t_{\rm gyro}$, given the observed rotation
periods, effective temperatures, and their uncertainties by
integrating Equation~1 of \citet{Bouma_2023} at the default grid
resolution.  This procedure marginalizes over the intrinsic age and
temperature dependent scatter that is observed in the cluster
sequences to yield statistical age uncertainties.

@@

\subsection{Gyrochronology Quality Flags}
\label{subsec:flags}
We calculated gyrochrone age posteriors for all
\nuniqstarsantosrotteffcut\ stars with reported rotation periods and
effective temperatures between 3800\,K and 6200\,K.  To flag stars for
which we suspect these ages may not in fact be valid, we built a set
of bit-quality flags.  When and how these flags should be applied
depends on the question being asked.  If the goal is to construct a
false-positive free sample, they might all be applied.  If the goal is
to construct a complete sample of young stars, then consider the
examples of Kepler~1627Ab ($\approx$40\,Myr) and Kepler~51c
($\approx$625\,Myr).  The former has a high RUWE due to a resolved
binary companion \citep{Bouma_2022a}; the latter is on a grazing orbit
\citep{2014ApJ...783...53M}.  Our gyrochronology quality flag, $Q_{\rm
gyro}$, enables selecting for or against such cases by selecting from
the following bits.

@@

{\it Surface gravity (bit 1)}---We flagged stars as possible subgiants
if they had $\log g < 4.2$.  We set this threshold by examining the
relevant flag from \citet{berger_2018_radii_evolnstates},
cross-matched against \citet{Berger_2020a_catalog}.

@@

An important decision for our subsequent analysis is how to combine
the above flags in order to decide whether or not a star is suited for
gyrochronology.  By default, we assume that a star is suitable for
gyrochronology if none of bits zero through nine (inclusive) are
raised.

\subsection{Planet Quality Flags}
\label{subsec:plflags}
For the subset of our analysis that focuses on the properties of
planets, we use the following additional quality indicators to assess 
the quality of a planet, $Q_{\rm planet}$.

% see Erik's 2022 discussion; might want to shift to include a cut on
% disposition score.
{\it Candidate reliability (bit 0)}---no false positives; only
``planet candidates'' and ``confirmed planets''.

{\it Candidate S/N (bit 1)}---the false positive rate increases
greatly toward Kepler's noise floor for planet detection (CITE).  We
required a S/N in excess of the usual floor, through a cut on the
``multiple event statistic'' (MES): we required ${\rm MES}$$>$10.

{\it Grazing planets (bit 2)}---grazing objects ($b$>0.8) often
provide non-robust planetary sizes due to the size-impact parameter
degeneracy.  Particularly for large planets, they can also include
astrophysical false positives at higher rates (due to the same
degeneracy).  While we include a flag for such objects, we by default
do not exclude them from our analysis; this is because...{\bf
justify}.

@@

\begin{figure}[!t]
	\begin{center}
		\leavevmode
		\includegraphics[width=0.49\textwidth]{li_vs_teff_koi_X_JUMP_eagles_logy.pdf}
	\end{center}
	\vspace{-0.6cm}
	\caption{
    {\bf Equivalent widths (EW) of the \ion{Li}{1} 6708\,\AA\ doublet
    for planet-hosting stars.} These measurements were made from
    Keck/HIRES spectra collected between 2009 and 2024.  Lines are the
    ``mean'' isochrones from the same models as in
    Figure~\ref{fig:models}; the intrinsic dispersion beyond
    $\approx$1\,Gyr becomes much larger than changes in the mean.  The
    logarithmic scale highlights the dynamic range of the detections
    and upper limits.  Some stars with lithium detections do not have
    detected rotation periods, and vice-versa.
		\label{fig:li_vs_teff}
	}
\end{figure}

Figure~\ref{fig:li_vs_teff} shows our measured lithium equivalent
widths, plotted over the mean isochrone models from
\texttt{EAGLES} \citep{Jeffries_2023}.  We declare an upper-limit for
plotting purposes if ${EW}-\sigma_{\rm EW}$$<$10\,m\AA.  However, our
EWs span -40 to 250\,m\AA; all negative EWs have uncertainties that
are statistically consistent with zero.

To calculate lithium ages, we use \texttt{EAGLES} (git commit
\texttt{ac09637}), which uses a similar method as
\texttt{gyro-interp}, and was calibrated on stars observed by the
Gaia-ESO spectroscopic survey in 52 open clusters with ages spanning 2
to 6{,}000\,Myr \citep{Jeffries_2023}.  We adopt a linear prior in
age, spanning $10^6$ to $10^{10.1}$\,yr, and symmetrize our EW
uncertainties for purposes of interfacing with \texttt{EAGLES} by
taking the maximum of the measured positive and negative 1$\sigma$
uncertainty intervals.
Certain quality flags described in Section~\ref{subsec:flags} are also
relevant for lithium ages, specifically those associated with
crowding and binarity since those effects can produce both systematic
and astrophysical biases in EW measurements.

@@

In terms of detected planet counts, our rotation-based ages for
the Kepler sample yield \{\nplyounggyro, \nplmidgyro, \nploldgyro\}
detected planets in the 0-1\,Gyr, 1-2\,Gyr, and 2-3\,Gyr bins.
For the sub-gigayear bin in particular, which holds specific weight in
studies of planet evolution, while there are \nplyounggyro\ planets
with median ages below 1\,Gyr, requiring $t_{\rm
gyro}$$<$1\,Gyr at 2$\sigma$ confidence yields
\nplyounggyrotwosigma\ planets orbiting \nplhostsyounggyrotwosigma\
stars.
The stellar and planetary properties of this latter sample of young
planets are given in Table~\ref{tab:planets}.

@@

\begin{figure*}[!t]
	\begin{center}
%		\subfloat{
%			\includegraphics[width=0.78\textwidth]{rp_vs_period_scatter_20240415_colorbyage_showaux-gyro_selsnrupper3_strictyoung.pdf}
%		}
%		
%		\subfloat{
			\includegraphics[width=0.78\textwidth]{rp_vs_period_scatter_20240415_colorbyage_showaux-gyro_anyyoung_anyyoung.pdf}
%		}
	\end{center}
	\vspace{-0.5cm}

	\caption{
		{\bf Sizes, orbital periods, and rotation-based ages of transiting exoplanets younger than one billion years.} 
		Planets highlighted in this plot have $t$<1\,Gyr at $2$$\sigma$.
%		
%		Two possible selection functions for what defines a young, sub-gigayear, planet
%		are imposed on the {\it top} and {\it bottom}: the former requires $t$<1\,Gyr with
%		$t/\sigma_{\rm t}$>3, the latter also allows planets younger than 1\,Gyr at 2$\sigma$.
%		Circles are previously reported measurements from the literature.
		Stars are newly calculated ages for Kepler planets from this work; circles
		are planets from the literature meeting the same requirement.
		Gray points are transiting planets from the NASA Exoplanet Archive older than
		1\,Gyr.
		The three age bins span 10-46\,Myr, 46-215\,Myr, and 215-1000\,Myr.
		\label{fig:rp_period_age_results}
	}
\end{figure*}

@@

\startlongtable
\begin{deluxetable*}{lllllllrrrrrrl}
	%\tabletypesize{\footnotesize}
	\tabletypesize{\scriptsize}
  \tablecaption{Ages of Kepler planets and planet candidates.  Rows
  are sorted by the minimum of either the rotation or lithium-based
  age, and ``confirmed'' planets appear before``candidate'' planets.
  This version of the table is truncated to planets younger than
  one billion years.  The full machine-readable table contains ages
  and age limits for \nnonfopkoissomeageinfo\ non-false positive KOIs
  with MES$>$10.  The bit quality flags for the rotation-based ages,
  $Q_{\rm gyro}$, are described in Section~\ref{subsec:flags}.
  Concisely summarized, they are:
  Bit 0: $T_{\rm eff}/{\rm K} \in [ 3800-6200]$ ?
  Bit 1: $\log g$$<$4.2?
  Bit 2: $M_{\rm G}$<3.9?
  Bit 3: In KEBC?
  Bit 4: Large $d_{\rm xm,Kep-Gaia}$?
  Bit 5: Confused Kep-Gaia crossmatch?
  Bit 6: Gaia DR3 non-single star?
  Bit 7: RUWE$>$1.4?
  Bit 8: Crowded?
  Bit 9: Far from main-sequence?
  Bit 10: \citetalias{Santos_2021} CP/CB?
  As an example, Kepler-1627 is flagged with bit 10 and bit 7.
  \label{tab:planets}}
	%\toprule
	%\midrule
	%\endhead
	\startdata
  KOI & Kepler &  $T_{\rm eff}$ & $P_{\rm rot}$ & EW$_{\rm Li}$ & $t_{\rm gyro}$ & $t_{\rm Li} $ & $R_{\rm p}$ & $P_{\rm orb}$ & $f_{\rm grazing}$ & $Q_{\rm gyro}$ &  Spec? & Comment \\
  -- &   -- & K & days & m\AA & Myr & Myr &    Earths &    days &       bool  & int & bool & -- \\
	\hline
	\input{table_subgyr_confirmed} %\\
	%\hline
	\input{table_subgyr_candidate}
	\enddata
%	\tablecomments{This table includes X, Y, Z... The machine-readable version,
%		available online, includes additional columns for .... 
%	}
\end{deluxetable*}

\startlongtable
\begin{deluxetable*}{lllllllrrrrrr}
	%\tabletypesize{\footnotesize}
	\tabletypesize{\scriptsize}
  \tablecaption{Ages of Kepler target stars.  
  The full machine-readable table contains rotation-based ages
  for \nuniqstarsantosrotteffcut\ stars with temperatures spanning
  3800-6200\,K.
  \nuniqstarsantosrotgyroappl\ of these stars have quality flags
  without bits 0 to 9 raised, meaning that they are likely valid
  objects for gyrochronology (see Section~\ref{subsec:flags}).
  \label{tab:stars}}
	%\toprule
	%\midrule
	%\endhead
	\startdata
  KIC & Gaia &  $T_{\rm eff}$ & $P_{\rm rot}$ & $t_{\rm gyro}$ & $Q_{\rm gyro}$ \\
  -- &   -- & K & days &  Myr &    int  \\
	\hline
  %FIXME make it!
	%\input{table_gyrostars} %\\
	%\hline
	\enddata
%	\tablecomments{This table includes X, Y, Z... The machine-readable version,
%		available online, includes additional columns for .... 
%	}
\end{deluxetable*}
